Tracing the cultural evolution of meaning at the semantics-pragmatics interface

### 2017-02-08 paper/new_simulation_setup.tex
results for larger state space + some code adjustments for plots and simulations

A lexicon $L$ is a mapping $\Messgs \rightarrow C$ from messages to concepts. With three
messages there are $6^3 = 343$ possible lexica, but many of these assign the same concept to
more than one message. For simplicity, we restrict attention to only those lexica which avoid
expressive redundancy. Hardcoding such an \emph{mutual exclusivity bias}
\citep[e.g.][]{Clark2009:Lexical-Meaning}, there are 20 non-redundant lexica. Functional
pressure towards efficient communication will evidently favor non-redundant lexica, so this
restriction is fairly innocuous but practical.

@@

\section*{Details}

@@

\begin{figure}[h!]
  \centering
  \includegraphics[scale=.35]{../code-LOT-extension/priors.png} %width = 0.75\linewidth
  \caption{Prior over types (lexica). The green prior corresponds to $t_{24}$.}
  \label{fig:priors}
\end{figure}

@@

\begin{figure}[h!]
  \centering
  \includegraphics[scale=.25]{../code-LOT-extension/u-l30-a1.png} %width = 0.75\linewidth
  \caption{Utility-Matrix for $\alpha = 1$ and $\lambda = 30$. The boxed value corresponds to $\text{EU}(t_{24},t_{24})$.}
  \label{fig:u-matrix}
\end{figure}

For $Q$ there are more values to test. For sequence length $k$ there's not that much change as long as it is not too small and possible noise due to the sampling of learning data is accounted for (it's still important, but I'll background this for now). As before, the posterior parameter $l$ is very important. If $l \geq 10$ we get very faithful transmission of our target type. Lower values lead to more mutation. Here's a visual impression of two $Q$ matrices for different $l$ values:
\begin{figure}[h!]
  \centering
  \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[scale=0.4]{../code-LOT-extension/q-l30-a1-k5-l1.png}
  \end{subfigure}
  ~
   \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[scale=0.4]{../code-LOT-extension/q-l30-a1-k5-l10.png}
  \end{subfigure}
  \caption{$Q$-matrices for $\alpha = 1, \lambda = 30, k = 5$, $200$ samples per parent type with either $l = 1$ (left) and $l = 10$ (right). The boxed value corresponds to $P(t_{24} | t_{24})$.}
  \label{fig:q-matrices}
\end{figure}

@@

\begin{figure}[h!]
  \centering
  \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[scale=0.4]{../code-LOT-extension/rmd-lam30-a1-k5-s200-l1.png}
  \end{subfigure}
  ~
   \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[scale=0.4]{../code-LOT-extension/rmd-lam30-a1-k5-s200-l10.png}
  \end{subfigure}
  \caption{Mean proportion of top $2/3$ types and target $t_{24}$ under replication only (green plot), mutation only (blue plot), and replication {\em and} mutation (red plot) after $50$ generations with $\alpha = 1, \lambda = 30, k = 5$, $200$ samples per parent type and $l = 1$ (left figure) and $l = 10$ (right figure).}
  \label{fig:outcomes}
\end{figure}

>>>>

A lexicon $L$ is a mapping $\Messgs \rightarrow C$ from messages to concepts. With three
messages there are $6^3 = 216$ possible lexica, but many of these assign the same concept to
more than one message. For simplicity, we restrict attention to only those lexica which avoid
expressive redundancy. Hardcoding such an \emph{mutual exclusivity bias}
\citep[e.g.][]{Clark2009:Lexical-Meaning}, there are 20 non-redundant lexica. Functional
pressure towards efficient communication will evidently favor non-redundant lexica, so this
restriction is fairly innocuous but practical.

@@

\section*{Details for mutual exclusivity type space}

@@

\begin{figure}[h!]
  \centering
  \includegraphics[scale=.35]{../code-LOT-extension/plots/priors-meTrue.png} %width = 0.75\linewidth
  \caption{Prior over types (lexica). The green prior corresponds to $t_{24}$.}
  \label{fig:priors}
\end{figure}

@@

\begin{figure}[h!]
  \centering
  \includegraphics[scale=.25]{../code-LOT-extension/plots/u-l30-a1-meTrue.png} %width = 0.75\linewidth
  \caption{Utility-Matrix for $\alpha = 1$ and $\lambda = 30$. The boxed value corresponds to $\text{EU}(t_{24},t_{24})$.}
  \label{fig:u-matrix}
\end{figure}

For $Q$ there are more values to test. For sequence length $k$ there's not that much change as long as it is not too small and possible noise due to the sampling of learning data is accounted for (it's still important, but I'll background this for now). As before, the posterior parameter $l$ is very important. If $l \geq 10$ we get very faithful transmission of our target type. Lower values lead to more mutation. Here's a visual impression of two $Q$ matrices for different $l$ values:
\begin{figure}[h!]
  \centering
  \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[scale=0.4]{../code-LOT-extension/plots/q-l30-a1-k5-l1-meTrue.png}
  \end{subfigure}
  ~
   \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[scale=0.4]{../code-LOT-extension/plots/q-l30-a1-k5-l10-meTrue.png}
  \end{subfigure}
  \caption{$Q$-matrices for $\alpha = 1, \lambda = 30, k = 5$, $200$ samples per parent type with either $l = 1$ (left) and $l = 10$ (right). The boxed value corresponds to $P(t_{24} | t_{24})$.}
  \label{fig:q-matrices}
\end{figure}

@@

\begin{figure}[h!]
  \centering
  \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[scale=0.4]{../code-LOT-extension/plots/rmd-lam30-a1-k5-s200-l1-meTrue.png}
  \end{subfigure}
  ~
   \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[scale=0.4]{../code-LOT-extension/plots/rmd-lam30-a1-k5-s200-l10-meTrue.png}
  \end{subfigure}
  \caption{Mean proportion of top $2/3$ types and target $t_{24}$ under replication only (green plot), mutation only (blue plot), and replication {\em and} mutation (red plot) after $50$ generations with $\alpha = 1, \lambda = 30, k = 5$, $200$ samples per parent type and $l = 1$ (left figure) and $l = 10$ (right figure).}
  \label{fig:outcomes}
\end{figure}

\section*{Details for larger type space}

Our target is now Gricean $t_{68}$ with the same lexicon as $t_{24}$ in the reduced type space. Following the setup above and the complexity values in Table \ref{tab:concepts}, the priors of our now $112$ types are shown here:

\begin{figure}[h!]
  \centering
  \includegraphics[scale=.35]{../code-LOT-extension/plots/priors-meFalse.png} %width = 0.75\linewidth
  \caption{Prior over types (lexica). The green prior corresponds to $t_{68}$.}
  \label{fig:priors}
\end{figure}

As before,  $t_{68}$ is not the best but also far from the worse. Here's a visualization of the entire u-matrix:

\begin{figure}[h!]
  \centering
  \includegraphics[scale=.25]{../code-LOT-extension/plots/u-l30-a1-meFalse.png} %width = 0.75\linewidth
  \caption{Utility-Matrix for $\alpha = 1$ and $\lambda = 30$. The boxed value corresponds to $\text{EU}(t_{68},t_{68})$.}
  \label{fig:u-matrix}
\end{figure}

Here's a visual impression of two $Q$ matrices for different $l$ values with the larger type space:

\begin{figure}[h!]
  \centering
  \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[scale=0.4]{../code-LOT-extension/plots/q-l30-a1-k5-l1-meFalse.png}
  \end{subfigure}
  ~
   \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[scale=0.4]{../code-LOT-extension/plots/q-l30-a1-k5-l10-meFalse.png}
  \end{subfigure}
  \caption{$Q$-matrices for $\alpha = 1, \lambda = 30, k = 5$, $200$ samples per parent type with either $l = 1$ (left) and $l = 10$ (right). The boxed value corresponds to $P(t_{68} | t_{68})$. You have to zoom in all the way to see something. Otherwise; it's similar to the mutation value of our target in the reduced space}
  \label{fig:q-matrices}
\end{figure}

Finally, we can see how the above affects actual applications of only replication, only mutation, and replication \emph{and} mutation using the matrices shown above. To construct these plots I took the mean top $3$ types from $1000$ simulations using either of the three dynamics. If $t_{68}$ was not in the top $3$, I took out the third most prolific type and added $t_{68}$ instead.

\begin{figure}[h!]
  \centering
  \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[scale=0.4]{../code-LOT-extension/plots/rmd-lam30-a1-k5-s200-l1-meFalse.png}
  \end{subfigure}
  ~
   \begin{subfigure}[b]{0.45\textwidth}
    \includegraphics[scale=0.4]{../code-LOT-extension/plots/rmd-lam30-a1-k5-s200-l10-meFalse.png}
  \end{subfigure}
  \caption{Mean proportion of top $2/3$ types and target $t_{68}$ under replication only (green plot), mutation only (blue plot), and replication {\em and} mutation (red plot) after $50$ generations with $\alpha = 1, \lambda = 30, k = 5$, $200$ samples per parent type and $l = 1$ (left figure) and $l = 10$ (right figure).}
  \label{fig:outcomes}
\end{figure}
